# Reproducible Research: Peer Assessment 1
## Ariel Israel, Jul 2013

## Loading and preprocessing the data
```{r read_data,echo=TRUE}
library(data.table)
data<-fread('d:/datasets/repdata/activity.csv')
```

## What is the sum of steps taken per day?
```{r number_of_steps_per_day,echo=TRUE}
steps_per_day<-data[,list(mean_steps=sum(steps,na.rm=T)),by='date']
hist(steps_per_day$mean_steps,nclass=50,col='#ff2500',xlab='steps per day',main='histogram of steps per day')
steps_summary=summary(steps_per_day$mean_steps)
```
* mean steps per day: `r format(steps_summary['Mean'],big.mark = ",")`
* median steps per day: `r format(steps_summary['Median'],big.mark = ",")`

## What is the average daily activity pattern?
```{r daily_activity_pattern,echo=TRUE}
steps_per_interval<-data[,list(mean_steps=sum(steps,na.rm=T)),by='interval']
plot(steps_per_interval$interval,steps_per_interval$mean_steps,xlab='steps per interval',main='time series steps per 5 minutes interval',type='l')
interval_max=with(steps_per_interval,interval[which.max(mean_steps)])
```
* 5 minutes interval containing the maximum number of steps: `r interval_max`


## Imputing missing values
* missing values in the dataset: `r sum(is.na(data$steps))`

## filling in with the missing values per day
Algorithm: For each row in which steps are unavailable, we use the median number of steps for this interval of the day:

- We do so by computing the median number of steps for each period of the day, and merging it with the original data under the column 'median_steps'
- Each time the steps column im the data in NA, we replace it with the median_steps value.
```{r imputing_missing_values,echo=TRUE}
# replace missing values with the median for the interval
data_imputed=merge(data,data[,list(median_steps=median(steps,na.rm=T)),by='interval'],by='interval')
to_fill_pos=is.na(data_imputed$steps)
data_imputed$steps[to_fill_pos]=data_imputed$median_steps[to_fill_pos]
data_imputed$median_steps=NULL
steps_per_day_imputed<-data_imputed[,list(mean_steps=sum(steps,na.rm=T)),by='date']
hist(steps_per_day_imputed$mean_steps,nclass=50,col='#ff2500',xlab='steps per day',main='histogram of steps per day')
steps_summary_imputed=summary(steps_per_day_imputed$mean_steps)
```
* missing values in the imputed dataset: `r sum(is.na(data_imputed$steps))`
* mean steps per day: `r format(steps_summary_imputed['Mean'],big.mark = ",")`
* median steps per day: `r format(steps_summary_imputed['Median'],big.mark = ",")`

The imputation step changed slightly the mean number of steps, but not the median

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday_differences,echo=TRUE}
data$weekday=weekdays(strptime(data$date,'%Y-%m-%d'),abbreviate=T)
data$weekday_or_weekend=factor(NA,levels=c('weekday','weekend'))
data$weekday_or_weekend[data$weekday %in% c('Mon','Tue','Wed','Thu','Fri')]='weekday'
data$weekday_or_weekend[data$weekday %in% c('Sat','Sun')]='weekend'
steps_per_interval_weekday<-data[,list(mean_steps=mean(steps,na.rm=T)),by='interval,weekday_or_weekend']
library(lattice)
xyplot(mean_steps~interval|weekday_or_weekend,data=steps_per_interval_weekday,type='l',layout=c(1,2))
```